Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-10 shanghai.tex
Frobenius fixes finished. I had to make some rather drastic notation changes.
I started by making the meaning of \Frob precise and consistent with
Deligne's definition of Weil sheaf. Deligne (and everyone else I could
find, like Laumon, Lusztig, Keihl and Weissauer, to name a few) uses F,
not Fr, for what we meant. So I had to change Fr to F in our paper. But
that made our current use of F, in the datum of a character sheaf, a
notational conflict. So I followed Lusztig in replacing F with \phi. I
also decided to harmonize our notation for trace of Frobenius with
Laumon, replacing our current t

\newcommand{\cdef}[1]{{{\color{blue}#1}\index{#1}}} 
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\obj}{{\text{obj}\, }}
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\newcommand{\GN}[1]{\mathswab{#1}}
\newcommand{\bGN}[1]{{\bar{\mathswab{#1}}}}
\newcommand{\TT}{\underline{T}}
\newcommand{\TL}{\underline{T_L}}
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\CS}{{\mathcal{C\hskip-1.8pt S}}}
\newcommand{\CSb}{{\mathcal{C\hskip-1.8pt S}_0}}
\newcommand{\CSf}{{\mathcal{C\hskip-1.8pt S}_{f}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\textit{iso}}}
\newcommand{\CSbiso}[1]{\CSb(#1)_{/\text{iso}}}
\newcommand{\CSfiso}[1]{\CSf(#1)_{/\text{iso}}}
\newcommand{\CE}{{\mathcal{C\hskip-1.8pt E}}}
\newcommand{\Lgroup}[1]{{\,^L\hskip-1pt{#1}}}
\newcommand{\dualgroup}[1]{{\check{#1}}}
\newcommand{\Lang}{{\operatorname{Lang}}}
\newcommand{\image}{{\operatorname{im}}}
\newcommand{\Weil}[1]{\mathcal{W}_{#1}}
\newcommand{\Loc}{{\operatorname{Loc}}}

@@

Since the character sheaves that appear in this paper are not quite an
instance of those defined by George Luzstig, who coined the term, we
begin by explaining what we mean by `character sheaves' here. 
Let $G$ be a smooth commutative group scheme locally of finite
type over $\Fq$; $G$ need not be reductive, connected, nor of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$. By a {character sheaf} on $G$ we mean a triple $\cs{L}\ceq
(\gcs{L},\mu,F)$ where $\gcs{L}$ is a constructible $\ell$-adic sheaf
$\gcs{L}$ on the \'etale site of $\bG \ceq G \times_{\Spec\Fq}
\Spec{\bFq}$, locally constant of rank-$1$ on every connected
component of $\bG$, equipped with an isomorphism $\mu: m^*
\gcs{L} \to \LxL$, where $m : \GxG \to G$ is
the group operation for $G$, and an isomorphism $F : \Frob{G}^*
\gcs{L} \to \gcs{L}$, where $\Frob{G}$ is the Frobenius automorphism
on $\bG$, satisfying natural compatibility conditions with $\mu$;
these compatibility conditions are explained below.

@@

Throughout this paper, $G$ stands for a commutative group scheme,
smooth and locally of finite type over a finite field $\Fq$; we write
$m : \GxG\to G$ for the multiplication morphism. We fix an
algebraic closure $\bFq$ of $\Fq$ and write $\bG $ for the
commutative group scheme, smooth and locally of finite type over
$\bFq$ obtained by base change; the multiplication morphism for
$\bG$ will be denoted by $\bm$. 
Let $\Frob{G}$ be the Frobenius automorphism of $\bG$.
\Clifton{I was a bit glib about Frobenius, previously.
Now I've made things more precise here; I'm in the process of tracking this improvement through the paper.}
Recall that $\Frob{G} = \id_{G}\times \Frob{\Spec{\bFq}}$ where
$\Frob{\Spec{\bFq}}$ is the automorphism of $\Spec{\bFq}$ 
opposite to the geometric Frobenius element $\Frob{\Fq}$ in $\Weil{\Fq} \subset \Gal(\bFq/\Fq)$.

The category of character sheaves on $G$, denoted by \cdef{$\CS(G)$},
is defined as follows.  A \cdef{character sheaf} on $G$ is a triple
$\cs{L}\ceq (\gcs{L},\mu,F)$ where:
\begin{enumerate}
\labitem{(CS.0)}{CS.0} $\gcs{L}$ is a rank-$1$ $\ell$-adic
  local system on (the \'etale site of) $\bG$;
\labitem{(CS.1)}{CS.1} $\mu: \bm^* \gcs{L} \to \LxL$
  is an isomorphism of sheaves on $\bGxG$ such that the following diagram commutes, 
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
%
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gcs{L} \boxtimes \gcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gcs{L} \boxtimes \bm^* \gcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gcs{L} \tight{0}{\boxtimes}{0} \LxL\\
  \end{tikzcd}
  \]

%
\labitem{(CS.2)}{CS.2} $F : \Frob{G}^* \gcs{L} \to \gcs{L}$ is an
  isomorphism of sheaves on $\bG$ compatible with $\mu$ in the sense that the following diagram commutes.
%
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gcs{L} \arrow[swap]{d}{\bm^* F}
    && \Frob{G}^*\gcs{L}\boxtimes \Frob{G}^*\gcs{L} \dar{F\boxtimes F} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL\\
  \end{tikzcd}
  \]
\end{enumerate}
Morphisms in $\CS(G)$ are defined in the natural way:
\begin{enumerate}
\labitem{(CS.3)}{CS.3} if $\cs{L} = (\gcs{L},\mu,F)$ and
  $\cs{L'} = (\gcs{L'},\mu',F')$ are character sheaves on $G$ then
  $\Hom_{\CS(G)}(\cs{L},\cs{L'})$ is the set of morphisms $\varphi : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gcs{L} \rar{\Frob{G}^* \varphi} \arrow[swap]{d}{F} & \Frob{G}^* \gcs{L'} \dar{F'}
  & \arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} & \arrow[swap]{d}{\mu} m^* \gcs{L} \rar{m^* \varphi} & m^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\varphi} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\varphi\boxtimes \varphi}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

@@

%\begin{proposition}\label{tensor}
The category $\CS(G)$ of character sheaves on $G$ is a rigid monoidal
category [insert reference] under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', F\otimes F')$
with duals given by applying the functor 
\Clifton{Should this be $\sheafHom(\ - \ ,\EE)$?} $\sheafHom(\ - \ ,(\EE)_{\bG})$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\CS(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\CS(G)$ puts a group
structure on the set $\CSiso{G}$ of isomorphism classes in
$\CS(G)$; we refer to this as the \cdef{Picard group of character
sheaves} on $G$.\footnote{See {\tt http://ncatlab.org/nlab/show/Picard+group}.}

@@

Let $\cs{L}$ be a character sheaf on $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob{G}$ and
therefore an automorphism $F_{\bg}$ of $\gcs{L}_{\bg}$. Note
that $\Aut_{\EE}(\gcs{L}_{\bg}) \iso \EEx$, canonically (this
follows from the Condition~\ref{CS.1}). The isomorphism $\mu : \bm^* \gcs{L} \to \LxL$
guarantees that the function $t^\Frob{\Fq}\cs{L} : G(\Fq)\to \EEx$,
defined by $g \mapsto F_{\bg}$, is a group homomorphism;
we refer to $t^\Frob{\Fq}\cs{L}$ as the \cdef{trace of Frobenius of $\cs{L}$}, 
even though no trace is involved in the definition, because the notion extends 
to any constructible $\ell$-adic Weil sheaf $(\gcs{L},F)$ on $G$, in which context,
$t^\Frob{(\gcs{L},F)} : G(\Fq)\to \EE$ is defined by $t^\Frob{(\gcs{L},F)}(g) \ceq \trace F_{\bg}$.

It is clear that $t^\Frob{\cs{L}\otimes \cs{L'}} = t^\Frob{\Fq}\cs{L} \cdot t^\Frob{\Fq}\cs{L'}$
(pointwise multiplication of functions) and that $t^\Frob{\Fq}\cs{L} = t^\Frob{\Fq}\cs{L'}$
if $\cs{L} \iso \cs{L'}$ in $\CS(G)$. In this way we obtain a group homomorphism
\[
\cdef{t_{G}} : \CSiso{G} \longrightarrow \Hom_\text{grp}(G(\Fq),\EEx),
\qquad \text{defined by}\qquad \cs{L} \mapsto t^\Frob{\Fq}\cs{L},
\]
from the Picard group of character sheaves on $G$ to the character group of $G(\Fq)$.

@@

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gcs{L},\mu,F) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \CS(G) \to \CS(H)$ such that
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \CSiso{G} \rar{f^*} \arrow[swap]{d}{t_{G}} & \CSiso{H} \dar{t_H} \\
  \Hom_\text{grp}(G(\Fq),\EEx) \rar & \Hom_\text{grp}(H(\Fq),\EEx)
  \end{tikzcd}
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

  To see that $(\bfxf)^* \gcs{L}$ satisfies
  condition \ref{CS.1} simply apply the functor $(\bfxf)^*$
  to \ref{CS.1} for $\cs{L}$ and use the canonical isomorphism
  $(\bfxf)^*(\LxL) \iso \bf^*\gcs{L} \tight{-3}{\boxtimes}{-3} \bf^*\gcs{L}$
  (a few times). Likewise, to see that $\bf^*\gcs{L}$ satisfies condition
  \ref{CS.2}, apply the same functor to \ref{CS.2} for $\gcs{L}$.
  Then use the fact that $f$ is defined over $\Fq$, so
  $(\bfxf)^*\Frob{\GxxG}^* \iso \Frob{\GxxG} (\bfxf)^*$,
  and the fact that $f$ is a morphism of group schemes, so
  $(\bfxf)^* \bm^*\iso \bm^* \bf^*$
  (isomorphisms of functors on constructible sheaves).

  To see that $(\gcs{L},\mu,F) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a functor $f^* : \CS(G) \to \CS(H)$ one applies the functors $\bf^*$ and
  $\bf^*\tight{1}{\times}{1}\bf^*$ to \ref{CS.3} and argue as
  above. It follows from the fact that sheaf hom commutes with these
  functors that $f^* : \CS(G) \to \CS(H)$ is a monoidal functor.

  The fact that the functor $f^* : \CS(G) \to \CS(H)$ commutes with
  the trace of Frobenius, in the sense above, follows immediately from
  the definitions; see \cite{Laumon}*{1.1.1.2}, where the ambient
  hypothesis that $X$ is of finite type over $\Fq$ can be replaced by
  the hypothesis that $X$ is locally of finite type over $\Fq$.

@@

When using character sheaves to study characters, it is useful to understand
how character sheaves behave under change of fields.
Suppose that $k'$ is a finite extension of $k$ of degree $n$, $G$ is an abelian group scheme
over $k$ as in Theorem \ref{thm:snake} and $G_{k'}$ is the base change of $G$ to $k'$.
If $\cs{L} = (\gcs{L}, \mu, F)$ is a character sheaf on $G$ we may define
a character sheaf $\nu(\cs{L}) = (\gcs{L}, \mu, F_{k'})$ on $G_{k'}$ by setting
\[
F_{k'} = F \circ \Frob{G}^*(F) \circ \cdots \circ (\Frob{G}^{n-1})^*(F).
\]
The commutativity of the diagram (CS.2) for $F_{k'}$ follows from the fact that $\Frob{G_{k'}} = \Frob{G}^n$.
Note that we may also think about the construction of $F_{k'}$ as taking the action of $\Weil{k}$ defined in Remark
\ref{rem:CS-action} and restricting it to $\Weil{k'}$.

\begin{proposition}\label{prop:basechange}
The following diagram commutes, where the bottom map is induced by the norm $G(k') \rightarrow G(k)$:
\[
\begin{tikzcd}
\CSiso{G} \rar{\nu} \dar{t_G} & \CSiso{G_{k'}} \dar{t_{G_{k'}}} \\
\Hom(G(k), \EEx) \rar{\Nm} & \Hom(G(k'), \EEx).
\end{tikzcd}
\]
\end{proposition}

@@

\begin{proof}
 Let $b_G : {\bar G} \to G$ be the pull-back of $\Spec{\bFq} \to \Spec{\Fq}$ along $G\to \Spec{\Fq}$.
 Let $(\cs{L}_0,\mu_0)$ be a bounded character sheaf on $G$. 
 Then $\cs{L}_0$ is, in particular, an $\ell$-adic constructible sheaf on $G$. 
 Then $b_G^* \cs{L}_0$ comes equipped with an isomorphism 
 $F : \Frob{G}^* b_G^*\cs{L}_0 \to b_G^* \cs{L}_0$; 
 moreover, the functor $\cs{L}_0 \mapsto (b_G^* \cs{L}_0,F)$  
 from $\ell$-adic constructible sheaves on $G$ to $\ell$-adic constructible sheaves on $G$
 is full and faithful; see \cite{SGA7.2}*{Expos\'e XIII} or even \cite{BBD}*{Prop. 5.2.1}. 
 This functor preserves local constancy, so takes local systems to local systems. 
 Set $\mu = b_{G\times G}^*\mu_0$; clearly, this satisfies Condition~\ref{CS.2} 
 with $b_G^*\cs{L}_0$ playing the role of $\gcs{L}$.
 Moreover, $F$ is compatible with $\mu$ in the sense of Condition~\ref{CS.3}.
 It follows that functor $(\cs{L}_0,\mu_0) \mapsto (b_G^*\cs{L}_0,b_{G\times G}^* \mu_0, F)$, denoted by  $B_G^*: \CSb(G) \hookrightarrow \CS(G)$ below, is full and faithful.
\end{proof}

@@

\section{Connected abelian algebraic groups}

@@

\begin{proposition}\label{prop:connected}
  If $G$ is a connected commutative algebraic group over $\Fq$ then 
  the full subcategories $\CSf(G) \hookrightarrow \CSb(G) \hookrightarrow \CS(G)$ 
  appearing in Section~\ref{sec:bounded} are equivalences 
  and (with reference to Section~\ref{sec:Frob})
  \[
  t_{G} : \CSiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx)
  \]
  is an isomorphism of groups.
\end{proposition}

>>>>

\newcommand{\cdef}[1]{{{\color{blue}#1}\index{#1}}} 
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\obj}{{\text{obj}\, }}
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\newcommand{\GN}[1]{\mathswab{#1}}
\newcommand{\bGN}[1]{{\bar{\mathswab{#1}}}}
\newcommand{\TT}{\underline{T}}
\newcommand{\TL}{\underline{T_L}}
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\CS}{{\mathcal{C\hskip-1.8pt S}}}
\newcommand{\CSb}{{\mathcal{C\hskip-1.8pt S}_0}}
\newcommand{\CSf}{{\mathcal{C\hskip-1.8pt S}_{f}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\textit{iso}}}
\newcommand{\CSbiso}[1]{\CSb(#1)_{/\text{iso}}}
\newcommand{\CSfiso}[1]{\CSf(#1)_{/\text{iso}}}
\newcommand{\CE}{{\mathcal{C\hskip-1.8pt E}}}
\newcommand{\Lgroup}[1]{{\,^L\hskip-1pt{#1}}}
\newcommand{\dualgroup}[1]{{\check{#1}}}
\newcommand{\Lang}{{\operatorname{Lang}}}
\newcommand{\image}{{\operatorname{im}}}
\newcommand{\Weil}[1]{\mathcal{W}_{#1}}
\newcommand{\Loc}{{\operatorname{Loc}}}
\newcommand{\trFrob}[1]{t_{#1}}

@@

Since the character sheaves that appear in this paper are not quite an
instance of those defined by George Luzstig, who coined the term, we
begin by explaining what we mean by `character sheaves' here. 
Let $G$ be a smooth commutative group scheme locally of finite
type over $\Fq$; $G$ need not be reductive, connected, nor of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$. By a {character sheaf} on $G$ we mean a triple $\cs{L}\ceq
(\gcs{L},\mu,\phi)$ where $\gcs{L}$ is a constructible $\ell$-adic sheaf
$\gcs{L}$ on the \'etale site of $\bG \ceq G \times_{\Spec\Fq}
\Spec{\bFq}$, locally constant of rank-$1$ on every connected
component of $\bG$, equipped with an isomorphism $\mu: m^*
\gcs{L} \to \LxL$, where $m : \GxG \to G$ is
the group operation for $G$, and an isomorphism $F : \Frob{G}^*
\gcs{L} \to \gcs{L}$, where $\Frob{G}$ is the Frobenius automorphism
on $\bG$, satisfying natural compatibility conditions with $\mu$;
these compatibility conditions are explained below.

@@

Throughout this paper, $G$ is a commutative group scheme,
smooth and locally of finite type over a finite field $\Fq$, and $m : \GxG\to G$ is its multiplication morphism.

We fix an algebraic closure $\bFq$ of $\Fq$ and write $\bG$ for the
commutative group scheme $G \times_{\Spec{\Fq}} \Spec{\bFq}$, over $\bFq$ obtained by base change from $k$. The multiplication morphism for $\bG$ will be denoted by $\bm$. 
%
Let $\Frob{G}$ be the Frobenius automorphism of $\bG$.
\Clifton{I was a bit glib about Frobenius, previously.
Now I've made things more precise here; I'm in the process of tracking this improvement through the paper.}
Recall that $\Frob{G} = \id_{G}\times \Frob{\Spec{\bFq}}$ where
$\Frob{\Spec{\bFq}} = \Spec\Frob{\Fq}$ and $\Frob{\Fq}$  is the geometric Frobenius element in $\Weil{\Fq} \subset \Gal(\bFq/\Fq)$.

The category of character sheaves on $G$, denoted by \cdef{$\CS(G)$},
is defined as follows.  A \cdef{character sheaf} on $G$ is a triple
$\cs{L}\ceq (\gcs{L},\mu,\phi)$ where:
\begin{enumerate}
\labitem{(CS.0)}{CS.0} $\gcs{L}$ is a rank-$1$ $\ell$-adic
  local system on (the \'etale site of) $\bG$;
\labitem{(CS.1)}{CS.1} $\mu: \bm^* \gcs{L} \to \LxL$
  is an isomorphism of sheaves on $\bGxG$ such that the following diagram commutes, 
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
%
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gcs{L} \boxtimes \gcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gcs{L} \boxtimes \bm^* \gcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gcs{L} \tight{0}{\boxtimes}{0} \LxL\\
  \end{tikzcd}
  \]

%
\labitem{(CS.2)}{CS.2} $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$ is an
  isomorphism of sheaves on $\bG$ compatible with $\mu$ in the sense that the following diagram commutes.
%
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gcs{L} \arrow[swap]{d}{\bm^* \phi}
    && \Frob{G}^*\gcs{L}\boxtimes \Frob{G}^*\gcs{L} \dar{F\boxtimes \phi} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL\\
  \end{tikzcd}
  \]
\end{enumerate}
Morphisms in $\CS(G)$ are defined in the natural way:
\begin{enumerate}
\labitem{(CS.3)}{CS.3} if $\cs{L} = (\gcs{L},\mu,\phi)$ and
  $\cs{L'} = (\gcs{L'},\mu',\phi')$ are character sheaves on $G$ then
  $\Hom_{\CS(G)}(\cs{L},\cs{L'})$ is the set of morphisms $\varphi : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gcs{L} \rar{\Frob{G}^* \varphi} \arrow[swap]{d}{F} & \Frob{G}^* \gcs{L'} \dar{F'}
  & \arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} & \arrow[swap]{d}{\mu} m^* \gcs{L} \rar{m^* \varphi} & m^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\varphi} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\varphi\boxtimes \varphi}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

@@

%\begin{proposition}\label{tensor}
The category $\CS(G)$ of character sheaves on $G$ is a rigid monoidal
category [insert reference] under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$
with duals given by applying the functor 
\Clifton{Should this be $\sheafHom(\ - \ ,\EE)$?} $\sheafHom(\ - \ ,(\EE)_{\bG})$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\CS(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\CS(G)$ puts a group
structure on the set $\CSiso{G}$ of isomorphism classes in
$\CS(G)$; we refer to this as the \cdef{Picard group of character
sheaves} on $G$.\footnote{See {\tt http://ncatlab.org/nlab/show/Picard+group}.}

@@

Let $\cs{L}$ be a character sheaf on $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob{G}$ and
therefore an automorphism $\phi_{\bg}$ of $\gcs{L}_{\bg}$. Note
that $\Aut_{\EE}(\gcs{L}_{\bg}) \iso \EEx$, canonically. The isomorphism $\mu : \bm^* \gcs{L} \to \LxL$
guarantees that the function $\trFrob{\cs{L}} : G(\Fq)\to \EEx$,
defined by $g \mapsto \phi_{\bg}$, is a group homomorphism;
we refer to $\trFrob{\cs{L}}$ as the \cdef{trace of Frobenius of $\cs{L}$}.
Although though no trace is involved in the definition, the notion extends 
to any constructible $\ell$-adic Weil sheaf $(\gcs{L},\phi)$ on $G$, in which context,
$\trFrob{(\gcs{L},\phi)} : G(\Fq)\to \EE$ is defined by $\trFrob{(\gcs{L},\phi)}(g) \ceq \trace \phi_{\bg}$.

It is clear that $\trFrob{\cs{L}\otimes \cs{L'}} = \trFrob{\cs{L}} \cdot \trFrob{\cs{L'}}$
(pointwise multiplication of functions) and that $\trFrob{\cs{L}} = \trFrob{\cs{L'}}$
if $\cs{L} \iso \cs{L'}$ in $\CS(G)$. In this way we obtain a group homomorphism
\[
\cdef{\trFrob{G}} : \CSiso{G} \longrightarrow \Hom_\text{grp}(G(\Fq),\EEx),
\qquad \text{defined by}\qquad \cs{L} \mapsto \trFrob{\cs{L}},
\]
from the Picard group of character sheaves on $G$ to the character group of $G(\Fq)$.

@@

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gcs{L},\mu,\phi) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \CS(G) \to \CS(H)$ such that
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \CSiso{G} \rar{f^*} \arrow[swap]{d}{\trFrob{G}} & \CSiso{H} \dar{t_H} \\
  \Hom_\text{grp}(G(\Fq),\EEx) \rar & \Hom_\text{grp}(H(\Fq),\EEx)
  \end{tikzcd}
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

  To see that $(\bfxf)^* \gcs{L}$ satisfies
  condition \ref{CS.1} simply apply the functor $(\bfxf)^*$
  to \ref{CS.1} for $\cs{L}$ and use the canonical isomorphism
  $(\bfxf)^*(\LxL) \iso \bf^*\gcs{L} \tight{-3}{\boxtimes}{-3} \bf^*\gcs{L}$
  (a few times). Likewise, to see that $\bf^*\gcs{L}$ satisfies condition
  \ref{CS.2}, apply the same functor to \ref{CS.2} for $\gcs{L}$.
  Then use the fact that $f$ is defined over $\Fq$, so
  $(\bfxf)^*\Frob{\GxxG}^* \iso \Frob{\GxxG}^* (\bfxf)^*$,
  and the fact that $f$ is a morphism of group schemes, so
  $(\bfxf)^* \bm^*\iso \bm^* \bf^*$
  (isomorphisms of functors on constructible sheaves).

  To see that $(\gcs{L},\mu,\phi) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a functor $f^* : \CS(G) \to \CS(H)$ one applies the functors $\bf^*$ and
  $\bf^*\tight{1}{\times}{1}\bf^*$ to \ref{CS.3} and argue as
  above. It follows from the fact that sheaf hom commutes with these
  functors that $f^* : \CS(G) \to \CS(H)$ is a monoidal functor.
%
  The fact that the functor $f^* : \CS(G) \to \CS(H)$ commutes with
  the trace of Frobenius, in the sense above, follows immediately from
  the definitions; see \cite{Laumon}*{1.1.1.2}, where the ambient
  hypothesis that $X$ is of finite type over $\Fq$ can be replaced by
  the hypothesis that $X$ is locally of finite type over $\Fq$.

@@

When using character sheaves to study characters, it is useful to understand
how character sheaves behave under change of fields.
Suppose that $k'$ is a finite extension of $k$ of degree $n$, $G$ is an abelian group scheme
over $k$ as in Theorem \ref{thm:snake} and $G_{k'}$ is the base change of $G$ to $k'$.
If $\cs{L} = (\gcs{L}, \mu, \phi)$ is a character sheaf on $G$ we may define
a character sheaf $\nu(\cs{L}) = (\gcs{L}, \mu, \phi_{k'})$ on $G_{k'}$ by setting
\[
\phi_{k'} = \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi).
\]
The commutativity of the diagram (CS.2) for $\phi_{k'}$ follows from the fact that $\Frob{G_{k'}} = \Frob{G}^n$.
Note that we may also think about the construction of $\phi_{k'}$ as taking the action of $\Weil{k}$ defined in Remark
\ref{rem:CS-action} and restricting it to $\Weil{k'}$.

\begin{proposition}\label{prop:basechange}
The following diagram commutes, where the bottom map is induced by the norm $G(k') \rightarrow G(k)$:
\[
\begin{tikzcd}
\CSiso{G} \rar{\nu} \dar{\trFrob{G}} & \CSiso{G_{k'}} \dar{\trFrob{G_{k'}}} \\
\Hom(G(k), \EEx) \rar{\Nm} & \Hom(G(k'), \EEx).
\end{tikzcd}
\]
\end{proposition}

@@

\begin{proof}
 Let $b_G : {\bar G} \to G$ be the pull-back of $\Spec{\bFq} \to \Spec{\Fq}$ along $G\to \Spec{\Fq}$.
 Let $(\cs{L}_0,\mu_0)$ be a bounded character sheaf on $G$. 
 Then $\cs{L}_0$ is, in particular, an $\ell$-adic constructible sheaf on $G$. 
 Then $b_G^* \cs{L}_0$ comes equipped with an isomorphism 
 $\phi : \Frob{G}^* b_G^*\cs{L}_0 \to b_G^* \cs{L}_0$; 
 moreover, the functor $\cs{L}_0 \mapsto (b_G^* \cs{L}_0,\phi)$  
 from $\ell$-adic constructible sheaves on $G$ to $\ell$-adic constructible sheaves on $G$
 is full and faithful; see \cite{SGA7.2}*{Expos\'e XIII} or even \cite{BBD}*{Prop. 5.2.1}. 
 This functor preserves local constancy, so takes local systems to local systems. 
 Set $\mu = b_{G\times G}^*\mu_0$; clearly, this satisfies Condition~\ref{CS.2} 
 with $b_G^*\cs{L}_0$ playing the role of $\gcs{L}$.
 Moreover, $\phi$ is compatible with $\mu$ in the sense of Condition~\ref{CS.3}.
 It follows that functor $(\cs{L}_0,\mu_0) \mapsto (b_G^*\cs{L}_0,b_{G\times G}^* \mu_0, \phi)$, denoted by  $B_G^*: \CSb(G) \hookrightarrow \CS(G)$ below, is full and faithful.
\end{proof}

@@

\section{Connected commutative algebraic groups}

@@

\begin{proposition}\label{prop:connected}
  If $G$ is a connected commutative algebraic group over $\Fq$ then 
  the full subcategories $\CSf(G) \hookrightarrow \CSb(G) \hookrightarrow \CS(G)$ 
  appearing in Section~\ref{sec:bounded} are equivalences 
  and (with reference to Section~\ref{sec:Frob})
  \[
  \trFrob{G} : \CSiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx)
  \]
  is an isomorphism of groups.
\end{proposition}

\begin{proof}
 %Let $G$ be a geometrically connected algebraic group over $\Fq$.
  Observe that the forgetful functor $(\gcs{L},\mu,\phi) \mapsto (\gcs{L},\phi)$
  sends character sheaves on $G$ to $\ell$-adic Weil sheaves on $G$ \cite{Deligne:Weil2}*{1.1.10}.
  While is it not true that all Weil sheaves on $G$ descend to local systems on $G$, 
  we will see that those that appear in the image of this forgetful functor from character sheaves do.
